{\bf Code Generation Log Replay for In-memory Database Management Systems}

### 2021-07-27 paper.tex
Updates from Overleaf

\begin{abstract}
Instead of interpreting log records, we are extending our query code generation engine to create custom programs to replay log records. We will also be exploring for different optimization techniques to improve the recovery speed of recovery with code generation.
\end{abstract}

@@

\subsection{Write-Ahead Logging}
\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{images/WAL.png}
\caption{Logging Architecture. Log manager uses the log serializer and log consumers to convert redo buffers into log records.}
\label{fig:wal}
\end{figure}

@@

\subsubsection{Physical to Logical Column Mapping}
The recovery manager needs to maintain a mapping from physical column ids to logical column ids for each table. This is required because the column ids recorded in the log records are physical ids, while all the query compilation components expect logical column ids. The mappings can be generated by looking into the DBMS catalog.
\subsubsection{Column Value Type Mapping}
For redo records, the recovery manager also needs to maintain a mapping from each column to its corresponding SQL value type (e.g. \dbSQL{INTEGER, FLOAT, VARLEN}). This mapping is used to correctly copy the values out of each redo record into a physical plan node.

@@

\section{Throughput Measurement}
\subsection{Insert}
We first measure the throughput (number of transactions executed per time period) of insert operations against the baseline recovery performance on a single table. The recovery manager uses only our insert implementation for recovery.
\begin{table}[H]
\begin{center}
\begin{tabular}{ |c|c|c| } 
 \hline
Statements Ratio & \textbf{Baseline Throughput} & \textbf{Codegen Throughput}\\ 
 \hline
 100\% Insert & $1.12 \times 10^4$ txns/s & $5.60 \times 10^3$ txns/s\\
 \hline
 50\% Insert, 50\% Select & $1.32 \times 10$ txns/s & $6.60 \times 10^3$ txns/s \\ 
 \hline
\end{tabular}
\caption{\textbf{Insert throughput measurement setup on a single table}. The table uses an initial table size of $10^6$. Each transaction runs 5 statements. A total of $10^5$ transactions are run.}
\label{tab:throughput_exp_insert}
\end{center}
\end{table}

2x difference. Memory copies.

@@

@@

\section{Throughput Measurement With Indexes}
In the baseline recovery architecture, the recovery manager implements a custom index update function that is independent from the query execution engine. This index update function is largely hard-coded and contains large amounts of memory copy operations. The code generation approach, however, use the built-in execution engine for index updates. We believe as the number of indexes increases, the performance gap between baseline and the code generation approach will shorten.
\begin{table}[H]
\begin{center}
\begin{tabular}{ |c|c|c|c|c| } 
 \hline
Statements Ratio & \#Indexes & \textbf{Baseline Throughput} & \textbf{Codegen Throughput} & \textbf{Baseline:Codegen} \\ 
 \hline
 100\% Insert & 5 & $3.7 \times 10^4$ txns/s & $2.4 \times 10^4$ txns/s & 1.54\\
 \hline
 100\% Insert & 10 & $2.4 \times 10^4$ txns/s & $1.7 \times 10^4$ txns/s & 1.41\\
 \hline
 100\% Insert & 100 & $3.2 \times 10^3$ txns/s & $2.4 \times 10^3$ txns/s & 1.3\\
 \hline
\end{tabular}
\caption{\textbf{Insert throughput measurement setup on a single table with indexes}. The table uses the same setup in \ref{tab:throughput_exp_insert}.}
\label{tab:throughput_exp_index_insert}
\end{center}
\end{table}

>>>>

\begin{abstract}
Code generation is a widely-used technique for improving query execution throughput by compiling instructions into native code. This technique, however, leads to design challenges for the recovery system of a database management system. The log replay process will be disjoint from the built-in execution engine that has been adapted to operate efficiently on compiled code. This usually leads to the implementation of a separate "execution engine" to deal with the execution of log records, which can be a huge waste of engineering efforts. To resolve this design conflict between code generation and database recovery, we propose a code generation design for the recovery system. We ask the recovery system to convert log records into compiled code that will be easily accepted by the execution engine. Our results show that while our approach creates more performance overhead compared to using a separate execution engine, it also introduces much less engineering overhead and is superior in aspects such as index updates and constraint checking.
\end{abstract}

@@

\subsection{Write-Ahead Logging}
\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{images/WAL.png}
\caption{\textbf{Logging Architecture}. Log manager uses the log serializer and log consumers to convert redo buffers into log records.}
\label{fig:wal}
\end{figure}

@@

\subsubsection{Physical to Logical Column Mapping}
The recovery manager needs to maintain a mapping from physical column ids to logical column ids for each table. This is required because the column ids recorded in the log records are physical ids, while all the query compilation components expect logical column ids. The mappings can be generated by looking into the DBMS catalog.
\subsubsection{Column Value Type Mapping}
For redo records, the recovery manager also needs to maintain a mapping from each column to its corresponding SQL value type (e.g. \dbSQL{INTEGER, FLOAT, VARLEN}). This mapping is used to correctly copy the values out of each redo record into a physical plan node.
\subsubsection{Catalog Access}
Retrieving access to a database catalog is costly. Profiling results show that a single operation to retrieve catalog access can account for 10\% of total execution time. Therefore, a mapping is used to keep track of catalog accessors.

@@

\section{Throughput Measurement}
\subsection{Insert}
\begin{table}[H]
\begin{center}
\begin{tabular}{ |c|c|c| } 
 \hline
Statements Ratio & \textbf{Baseline Throughput} & \textbf{Codegen Throughput}\\ 
 \hline
 100\% Insert & $1.12 \times 10^4$ txns/s & $6.00 \times 10^3$ txns/s\\
 \hline
 50\% Insert, 50\% Select & $1.32 \times 10$ txns/s & $6.60 \times 10^3$ txns/s \\ 
 \hline
\end{tabular}
\caption{\textbf{Insert throughput measurement setup on a single table}. The table uses an initial table size of $10^6$. Each transaction runs 5 statements. A total of $10^5$ transactions are run.}
\label{tab:throughput_exp_insert}
\end{center}
\end{table}
We first measure the throughput (number of transactions executed per time period) of insert operations against the baseline recovery performance on a single table. The recovery manager uses only our insert implementation for recovery.

\begin{table}[H]
\begin{center}
\begin{tabular}{ |c|c|c| } 
 \hline
\textbf{IR} & \textbf{Percentage} & \textbf{Functionality} \\ 
 \hline
79.35 & \textbf{81.7\%} & Query Execution (Op Codes)\\
 \hline
 9.35 & 9.62\% & Execution Context Initialization \\ 
 \hline
 6.19 & 6.37\% & Retrieving Redo Values \\
 \hline
 2.26 & 2.32 \% & Metadata  Cache Map Access \\
 \hline
 0.03 & 0.03 \% & Compile \\
 \hline
\end{tabular}

\begin{tabular}{ |c|c|c| } 
 \hline
\textbf{IR} & \textbf{Percentage} & \textbf{Operation Code}\\ 
 \hline
29.88 & \textbf{74.4\%} & \texttt{StorageInterfaceInit}\\
 \hline
 6.00 & 14.9\% & \texttt{StorageInterfaceTableInsert} \\ 
 \hline
 2.76 & 6.87\% & \texttt{StorageInterfaceFree} \\
 \hline
 1.40 & 3.48\% & \texttt{StorageInterfaceGetTablePR} \\
 \hline
 0.12 & 0.35\% & \texttt{ExecOUFeatureVectorReset} \\
 \hline
\end{tabular}
\caption{\textbf{KCachegrind Analysis on Insert}. IR refers to the number of instructions executed in total by the selected function during a single recovery step.}
\label{tab:throughput_exp_insert_overhead_all}
\end{center}
\end{table}
Results show that the code generation throughput is 47\% slower than the baseline implementation. Profiling result \ref{tab:throughput_exp_insert_overhead_all} explains that 81.7\% of the slow down comes from the execution of the compiled query. This is unexpected, as we imangined that most of the overhead would come from memory copies for extracting values, but those memory copy overhead only accounts for 6\%.

A further dive into the composition of query execution shows that 74.4\% of the overhead comes from \texttt{StorageInterfaceInit}, which is a function that initializes a storage interface. The storage interface stores objects that allow manipulation over a certain DBMS table. These objects can include the execution context, redo deltas and index objects associated with table. The storage interface is reconstructed every time an execution of a query begins, and is destructed after the query has been executed.

@@

Performance of delete is closer to the baseline than insert, with code generation performance aroudn 90\% to 95\% of baseline performance. The performance of delete is close to our expectation, as delete does not require any memory copies from a redo record like an insert does. Executing a delete is also a lightweight operation, as we have prevented the pipeline from performing extra work as shown in Section 3.2.2.

@@

\section{Throughput Measurement With Indexes}
In the baseline recovery architecture, the recovery manager implements a custom index update function that is independent from the query execution engine. This index update function is largely hard-coded and contains large amounts of memory copy operations. The code generation approach, however, use the built-in execution engine for index updates. We believe as the number of indexes increases, the performance gap between baseline and the code generation approach will shorten.
\begin{table}[H]
\begin{center}
\begin{tabular}{ |c|c|c|c|c| } 
 \hline
Statements Ratio & \#Indexes & \textbf{Baseline Throughput} & \textbf{Codegen Throughput} & \textbf{Baseline:Codegen} \\ 
 \hline
 100\% Insert & 5 & $3.97 \times 10^4$ txns/s & $2.65 \times 10^4$ txns/s & 1.50\\
 \hline
 100\% Insert & 10 & $2.37 \times 10^4$ txns/s & $1.7 \times 10^4$ txns/s & 1.40\\
 \hline
 100\% Insert & 100 & $3.30 \times 10^3$ txns/s & $2.49 \times 10^3$ txns/s & 1.32\\
 \hline
\end{tabular}
\caption{\textbf{Insert throughput measurement setup on a single table with indexes}. The table uses the same setup in \ref{tab:throughput_exp_insert}.}
\label{tab:throughput_exp_index_insert}
\end{center}
\end{table}
